---
title: "shinydevcon notes"
author: "Mika Braginsky"
output: 
  html_document: 
    highlight: tango
    theme: spacelab
    toc: yes
---

# Reactivity tutorial

> Keep your side effects
> 
> Outside of your reactives
> 
> Or I will kill you
> 
> -- Joe Cheng

[Talk](https://rawgit.com/rstudio/reactivity-tutorial/master/slides.html)

Takeaways for Wordbank improvement:

- make sure there are no side-effects in reactives (esp modifying UI)
    - `reactive()` is for calculating values, without side effects
    - `observe()` is for performing actions, with side effects
- probably use reactiveValues for stuff, it's useful
- use `req()` in reactive to depend on non-null input values (should help reactive UI elements resolve way more reasonably)
    - THIS WILL FIX EVERYTHING!!!
- use `shinyReactLog` to investigate input/reactive/output dependencies and app execution flow
- use `shinyURL` to share apps state by url (e.g. here's the trajectory for _dog_)
    - lmwtfy: let me wordbank that for you!
- `shinySignals` might be a thing?


# Plot interactivity

[Talk](https://gist.github.com/wch/25a1c4ce8fc84022d3e7),
[Article](http://shiny.rstudio.com/articles/plot-interaction.html)

Get click/dblclick/hover/brush events by coordinates (can use to retrieve nearest points, select/add/delete points, zoom in on plot region).


# Gadgets

[Article about gadgets](http://shiny.rstudio.com/articles/gadgets.html), [article about gadgets UI](http://shiny.rstudio.com/articles/gadget-ui.html)

Differences between a shiny app (for sharing things on the web) and a shiny gadget (for doing things in RStudio):

- call shiny gadget like a function
- displayed with miniUI elements
- always returns a value

Example: make scatterplot, select points to exclude from analysis

Furthermore, you can make a gadget into an RStudio add-in, which gives you a modal dialogue with the gadget and then replaces the selection in IDE with the results when done ([like so](https://github.com/hadley/hadladdin) and [more generally so](https://rstudio.github.io/rstudioaddins/)).

Let's you do things that are hard to express with code but want to do reproducibly.

Let's combine with plot interactivity to make an awesome ggplot fiddler!!! :D :D :D


# Deploying

Coming soon: seemless reconnection (momentary network blips won't make apps grey out, yaaay). Will be available on both shiny server and shinyapps.io, with a configurable timeout (default 15s).

New product: Rstudio Connect

  - hub for sharing shiny, Rmd, html, plots
  - push button publish from IDE
  - dashboard with all published things
  - specify who can access and publish
  - Rmarkdown reports aren't just static html, can re-run (manually and scheduled)
  - easily accesible log files
  
This looks absolutely fantastic! [Beta here](beta.rstudioconnect.com).

Parametrized Rmarkdown documents: takes a bunch of arguments, then knit with parameters (this should be great to use for things like local vs. remote wordbank connection).


# Modules

[Article](http://shiny.rstudio.com/articles/modules.html)

- Quickly resuse the same code in different apps or multiple times in the same app
- Divide code into separate modules that can be reasoned about independently
- Structure
    - write function that returns shiny UI (wrap multiple elements in `tagList()`)
    - assign module elements to a unique namespace with `NS()`
    - write server function
    - load server module with app's server with `callModule()`
    - module functions can be defined in:
        - single file app preamble
        - file sourced in single file app preamble
        - in `global.R`
        - in a file sourced by `global.R`
        - in a package that the app loads
    
Minimal example:    

```{r, eval=FALSE}
library(shiny)

sliderTextUI <- function(id) {
  ns <- NS(id)
  tagList(
    sliderInput(ns("slider"), "Slide me", 0, 100, 1),
    textOutput(ns("number"))
  )
}

sliderText <- function(input, output, session) {
  output$number <- renderText({
    input$slider
  })
}

ui <- fluidPage(
  sliderTextUI("one"),
  sliderTextUI("two")
)

server <- function(input, output) {
  callModule(sliderText, "one")
  callModule(sliderText, "two")
}

shinyApp(ui, server)
```

Use reactive expressions to pass reactive information between functions. Makes modules interact with other app components, not just a set functions.

  - Reactive output:
    - in server module function, return reactive output as a [list of] reactive expression[s]
    - `callModule()` returns all the output returns by server module function
  - Reactive input:
    - in server module function, wrap input as a reactive expression
    - pass the reactive expression to the modules as an argument
    - treat the argument as a reactive expression within the function
  
Not a ton of re-use between Wordbank apps, but might be able to factor out language/form/measure selection into a module.


# Debugging

[Talk](http://www.github.com/jmcphers/shiny-debug-talk), [Article](http://shiny.rstudio.com/articles/debugging.html)

Debugging:

- Get into debugger: `browser()`
- Go into showcase mode: `runApp("MyApp", display.mode = "showcase")`
- Look at reactive log (or `Cmd + F3`)
```{r, eval = FALSE}
options(shiny.reactlog = TRUE)
runApp(...)
showReactLog()
```
- Print things to stderr: `cat("hello world", file = stderr())` (watch out for this changing the reactivity structure and making Heisenbugs)
- Trace on shinyapps.io `rsconnect::showLogs(streaming = TRUE)`
- Trace on a shiny server `tail -f /var/log/shiny-server/myapp-*.log`
- Trace client/server messages `options(shiny.trace = TRUE)`

Error handling:

  - Tracebacks (turn on full with `options(shiny.fullstacktrace = TRUE)`)
  - Pause on errors with `options(shiny.error = browser)`
  - Look at JS errors in developer tools in browser
  - Turn on JS dev mode on OS X in RStudio `defaults write org.rstudio.RStudio WebKitDeveloperExtras -bool true`


# UI

[Article about templates](http://shiny.rstudio.com/articles/templates.html)

- Use R function to HTML and Shiny components in `ui.R`
- Use HTML to write HTML and Shiny components in `index.html`
- Use HTML to write HTML components and R to write Shiny components with `htmlTemplates`
    - write template in HTML
    - insert R code between `{{ }}`
    - call with htmlTemplate() in ui
    - add css, scripts, other web dependencies, etc
    - can define entire webpage or components to include


# Dashboards

- `reactiveFileReader` gives a reactive data source for the contents of a file
- `reactivePoll` gives a reactive data source using a checkFunc (equality check to see if values have changed) and valueFunc (retrieves values)
- `shinydashboard` package formats UI with specific nice structure
- `htmlTemplate` can then apply custom look to dashboard elements


# Profiling

- Use `RProf` with `profvis` to profile and visualize code execution ([vignettes](http://rpubs.com/wch/123888))
- Wrap whole shiny app in `profvis` call to profile app (all executed code must be within call)


# Data tables

Create a [htmlwidgets](http://www.htmlwidgets.org) to use a JS visualization in R; can then use widget in console, markdown, shiny.


# shinyjs

[GitHub](https://github.com/daattali/shinyjs), [slides](https://docs.google.com/presentation/d/1XyBrIA97ZWIkm8NICAyWHo4J9hj_i2-mJmKEq_CR2A8/edit#slide=id.p), [examples](https://gist.github.com/daattali/b95fe2376c9e16f35f15)
